---
title: "Household Pulse Food Calculations"
output: html_document
---

#THIS CODE RUNS FOR WEEKS 1-12 ONLY THE 4-WEEK ROLLING AVERAGE, FOR FOOD INSUFFICIENCY PERCENTAGE
## This code uses data imported from Food Tables 2, 3, and 5 in order to calculate percentages used in the data viz.



```{r}
library(tidyverse)
library(readxl)
library(httr)
library(stringr)
```

## STEP 1: UPDATE THE MOST RECENT WEEK FOR DOWNLOADING DATA
```{r}
week <- 12
state_abbr <- 'SD'
```


## STEP 2: CALCULATE ROLLING 4 "WEEK" AVERAGES
# 2a) For Sometimes/Often Not Enough::Overalll, Households with Children, Households without Children
```{r}
week_minus1 <- week-1
week_minus2 <- week-2
week_minus3 <- week-3

####### -- 2A) Calculate the Sometimes/Often Not Enough Percentages First -- #####

#sql code to import most recent 4 weeks of food 2b data
food_sql <- paste0("SELECT * FROM food2b WHERE state='",state_abbr,"' AND (week='",week,"' OR week='",week_minus1,"' OR week='",week_minus2,"' OR week='",week_minus3,"');")

food_4weeks_data <- dbGetQuery(con,food_sql) %>%
  
  subset(categorynames=='children' | categorynames=='total') %>%
  
  #covert variables to numeric
  mutate(foodinsufficient_numerator=as.numeric(paste(foodinsufficient_numerator)),
         foodinsufficient_numerator_se=as.numeric(paste(foodinsufficient_numerator_se)),
         totalvalidresponses=as.numeric(paste(totalvalidresponses)),
         totalvalidresponses_se=as.numeric(paste(totalvalidresponses_se))) %>%
  
  #combine four weeks of data
  group_by(select_characteristics,categorynames,state) %>%
  summarise(foodinsufficient_numerator=sum(foodinsufficient_numerator,na.rm=TRUE),
            foodinsufficient_numerator_se=sqrt((sum(foodinsufficient_numerator_se^2))),
            totalvalidresponses=sum(totalvalidresponses,na.rm=TRUE),
            totalvalidresponses_se=sqrt((sum(totalvalidresponses_se^2))),
            .groups='keep') %>%
  ungroup %>%
  arrange(categorynames,select_characteristics) %>%
  
  #calculate percent food insufficient, se, and confidence interval
  mutate(percent=foodinsufficient_numerator/totalvalidresponses) %>%
  mutate(percent_se=(1/totalvalidresponses)*(sqrt((foodinsufficient_numerator_se^2)-((percent^2)*(totalvalidresponses_se^2))))) %>%
  mutate(percent_se=if_else(is.na(percent_se),(1/totalvalidresponses)*(sqrt((foodinsufficient_numerator_se^2)+((percent^2)*(totalvalidresponses_se^2)))),percent_se)) %>%
  mutate(percent_lci=percent-(percent_se*1.645)) %>%
  mutate(percent_uci=percent+(percent_se*1.645)) %>%
  
  #lower confidence interval shouldn't be less than zero, and upper more than 100
  mutate(percent_lci=replace(percent_lci,percent_lci<0,0)) %>%
  mutate(percent_uci=replace(percent_uci,percent_uci>1,1)) %>%
  
  
  #add indicator variable for unreliable estimates
  mutate(percent_relativese=percent_se/percent) %>%
  mutate(keep=if_else(percent_relativese>.3,0,1)) %>%
  
  #remove unneeded variables
  select(-c(foodinsufficient_numerator,foodinsufficient_numerator_se,
            totalvalidresponses,totalvalidresponses_se,percent_relativese)) %>%
  
  #add back in database variables
  mutate(week_start=all_of(week_minus3),
         week_end=all_of(week),
         timeframe='4-week rolling average',
         varname='sometimesoftennotenoughtoeat')





#DATA QUALITY CHECKS
View(food_4weeks_data)
```


## STEP 3: ADD ROLLING AVERAGE DATA TO SQL DATABASE
```{r}
dbWriteTable(con,'calculatedpercents',food_4weeks_data,append=TRUE,row.names=FALSE)
```


