---
title: "Household Pulse Food Calculations"
output: html_document
---

## This code uses data imported from Food Tables 2, 3, and 5 in order to calculate percentages used in the data viz.



```{r}
library(tidyverse)
library(readxl)
library(httr)
library(stringr)
```

## STEP 1: UPDATE THE MOST RECENT WEEK FOR DOWNLOADING DATA
```{r}
week <- 27
state_abbr <- 'SD'
```


## STEP 2: CALCULATE ROLLING 4 "WEEK" AVERAGES
# 2a) For Sometimes/Often Not Enough::Overalll, Households with Children, Households without Children
# 2b) For Free Groceries/Meal in Last 7 Days:: Overall, Households with Children
# 2c) For Children Not Eating Because Food was Unaffordable:: Households with Children
# 2d) Combine data tables from 2a)-2c)
```{r}
week_minus1 <- week-1
week_minus2 <- week-2
week_minus3 <- week-3

####### -- 2A) Calculate the Sometimes/Often Not Enough Percentages First -- #####

#sql code to import most recent 4 weeks of food 2b data
food_sql <- paste0("SELECT * FROM food2b WHERE state='",state_abbr,"' AND (week='",week,"' OR week='",week_minus1,"' OR week='",week_minus2,"' OR week='",week_minus3,"');")

food_4weeks_data <- dbGetQuery(con,food_sql) %>%
  
  subset(categorynames=='children' | categorynames=='total') %>%
  
  #covert variables to numeric
  mutate(foodinsufficient_numerator=as.numeric(paste(foodinsufficient_numerator)),
         foodinsufficient_numerator_se=as.numeric(paste(foodinsufficient_numerator_se)),
         totalvalidresponses=as.numeric(paste(totalvalidresponses)),
         totalvalidresponses_se=as.numeric(paste(totalvalidresponses_se))) %>%
  
  #combine four weeks of data
  group_by(select_characteristics,categorynames,state) %>%
  summarise(foodinsufficient_numerator=sum(foodinsufficient_numerator,na.rm=TRUE),
            foodinsufficient_numerator_se=sqrt((sum(foodinsufficient_numerator_se^2))),
            totalvalidresponses=sum(totalvalidresponses,na.rm=TRUE),
            totalvalidresponses_se=sqrt((sum(totalvalidresponses_se^2))),
            .groups='keep') %>%
  ungroup %>%
  arrange(categorynames,select_characteristics) %>%
  
  #calculate percent food insufficient, se, and confidence interval
  mutate(percent=foodinsufficient_numerator/totalvalidresponses) %>%
  mutate(percent_se=(1/totalvalidresponses)*(sqrt((foodinsufficient_numerator_se^2)-((percent^2)*(totalvalidresponses_se^2))))) %>%
  mutate(percent_se=if_else(is.na(percent_se),(1/totalvalidresponses)*(sqrt((foodinsufficient_numerator_se^2)+((percent^2)*(totalvalidresponses_se^2)))),percent_se)) %>%
  mutate(percent_lci=percent-(percent_se*1.645)) %>%
  mutate(percent_uci=percent+(percent_se*1.645)) %>%
  
  #lower confidence interval shouldn't be less than zero, and upper more than 100
  mutate(percent_lci=replace(percent_lci,percent_lci<0,0)) %>%
  mutate(percent_uci=replace(percent_uci,percent_uci>1,1)) %>%
  
  
  #add indicator variable for unreliable estimates
  mutate(percent_relativese=percent_se/percent) %>%
  mutate(keep=if_else(percent_relativese>.3,0,1)) %>%
  
  #remove unneeded variables
  select(-c(foodinsufficient_numerator,foodinsufficient_numerator_se,
            totalvalidresponses,totalvalidresponses_se,percent_relativese)) %>%
  
  #add back in database variables
  mutate(week_start=all_of(week_minus3),
         week_end=all_of(week),
         timeframe='4-week rolling average',
         varname='sometimesoftennotenoughtoeat')


####### -- 2B) Calculate the Free Groceries Percentages Next -- #####
#In order to have both total and households with children, need data from tables 2b and 3b

food_4weeks_data_overallfree <- dbGetQuery(con,food_sql) %>%
  
  subset(categorynames=='freegroceries') %>%
  subset(select_characteristics != "Did not report") %>%
  
  #covert variables to numeric
  mutate(total=as.numeric(paste(total)),
         total_se=as.numeric(paste(total_se)),
         enough_types_food=as.numeric(paste(enough_types_food)),
         enough_types_food_se=as.numeric(paste(enough_types_food_se)),
         enough_not_types=as.numeric(paste(enough_not_types)),
         enough_not_types_se=as.numeric(paste(enough_not_types_se)),
         sometimes_not_enough=as.numeric(paste(sometimes_not_enough)),
         sometimes_not_enough_se=as.numeric(paste(sometimes_not_enough_se)),
         often_not_enough=as.numeric(paste(often_not_enough)),
         often_not_enough_se=as.numeric(paste(often_not_enough_se))) %>%
  
  #combine four weeks of data
  group_by(select_characteristics,categorynames,state) %>%
  summarise(count_sums=sum(enough_types_food,enough_not_types,sometimes_not_enough,often_not_enough,na.rm=TRUE),
            count_sums_se=sqrt((sum(enough_types_food_se^2)+sum(enough_not_types_se^2)+sum(sometimes_not_enough_se^2)+(sum(often_not_enough_se^2)))),.groups='keep')

#create an overall column to merge
overall_sum <- food_4weeks_data_overallfree %>%
  group_by(categorynames,state) %>%
  summarise(denominator_overall=sum(count_sums,na.rm=TRUE),
            denominator_overall_se=sqrt(sum(count_sums_se^2)),.groups='keep') 

#merge overall back with Yes/No
food_4weeks_data_overallfree2 <- food_4weeks_data_overallfree %>%
  left_join(overall_sum,by=c('categorynames'='categorynames','state'='state')) %>%
  
  #calculate percentages, keep only Yes
  subset(select_characteristics=='Yes') %>%
  mutate(percent=count_sums/denominator_overall,
         percent_se=(1/denominator_overall)*(sqrt((count_sums_se^2)-((percent^2)*(denominator_overall_se^2))))) %>%
  mutate(percent_se=if_else(is.na(percent_se),(1/denominator_overall)*(sqrt((count_sums_se^2)+((percent^2)*(denominator_overall_se^2)))),percent_se)) %>%
  mutate(percent_lci=percent-(percent_se*1.645)) %>%
  mutate(percent_uci=percent+(percent_se*1.645)) %>%
  
  #lower confidence interval shouldn't be less than zero, and upper more than 100
  mutate(percent_lci=replace(percent_lci,percent_lci<0,0)) %>%
  mutate(percent_uci=replace(percent_uci,percent_uci>1,1)) %>%
  
  
  #add indicator variable for unreliable estimates
  mutate(percent_relativese=percent_se/percent) %>%
  mutate(keep=if_else(percent_relativese>.3,0,1)) %>%
  
  #remove unneeded variables
  select(-c(count_sums,count_sums_se,
            denominator_overall,denominator_overall_se,percent_relativese)) %>%
  
  #add back in database variables
  mutate(week_start=all_of(week_minus3),
         week_end=all_of(week),
         timeframe='4-week rolling average',
         varname='freegroceriesorfreemeal',
         select_characteristics='Total')


### -- Repeat Steps for Table 3 -- ###
#sql code to import most recent 4 weeks of food 2b data
food_sql2 <- paste0("SELECT * FROM food3b WHERE state='",state_abbr,"' AND (week='",week,"' OR week='",week_minus1,"' OR week='",week_minus2,"' OR week='",week_minus3,"');")

food_4weeks_data_childfree <- dbGetQuery(con,food_sql2) %>%
  
  subset(categorynames=='freegroceries') %>%
  subset(select_characteristics != "Did not report") %>%
  
  #covert variables to numeric
  mutate(total=as.numeric(paste(total)),
         total_se=as.numeric(paste(total_se)),
         enough_types_food=as.numeric(paste(enough_types_food)),
         enough_types_food_se=as.numeric(paste(enough_types_food_se)),
         enough_not_types=as.numeric(paste(enough_not_types)),
         enough_not_types_se=as.numeric(paste(enough_not_types_se)),
         sometimes_not_enough=as.numeric(paste(sometimes_not_enough)),
         sometimes_not_enough_se=as.numeric(paste(sometimes_not_enough_se)),
         often_not_enough=as.numeric(paste(often_not_enough)),
         often_not_enough_se=as.numeric(paste(often_not_enough_se))) %>%
  
  #combine four weeks of data
  group_by(select_characteristics,categorynames,state) %>%
  summarise(count_sums=sum(enough_types_food,enough_not_types,sometimes_not_enough,often_not_enough,na.rm=TRUE),
            count_sums_se=sqrt((sum(enough_types_food_se^2)+sum(enough_not_types_se^2)+sum(sometimes_not_enough_se^2)+(sum(often_not_enough_se^2)))),.groups='keep')

#create an overall column to merge
overall_sum_child <- food_4weeks_data_childfree %>%
  group_by(categorynames,state) %>%
  summarise(denominator_overall=sum(count_sums,na.rm=TRUE),
            denominator_overall_se=sqrt(sum(count_sums_se^2)),.groups='keep') 

#merge overall back with Yes/No
food_4weeks_data_childfree2 <- food_4weeks_data_overallfree %>%
  left_join(overall_sum_child,by=c('categorynames'='categorynames','state'='state')) %>%
  
  #calculate percentages, keep only Yes
  subset(select_characteristics=='Yes') %>%
  mutate(percent=count_sums/denominator_overall,
         percent_se=(1/denominator_overall)*(sqrt((count_sums_se^2)-((percent^2)*(denominator_overall_se^2))))) %>%
  mutate(percent_se=if_else(is.na(percent_se),(1/denominator_overall)*(sqrt((count_sums_se^2)+((percent^2)*(denominator_overall_se^2)))),percent_se)) %>%
  mutate(percent_lci=percent-(percent_se*1.645)) %>%
  mutate(percent_uci=percent+(percent_se*1.645)) %>%
  
  #lower confidence interval shouldn't be less than zero, and upper more than 100
  mutate(percent_lci=replace(percent_lci,percent_lci<0,0)) %>%
  mutate(percent_uci=replace(percent_uci,percent_uci>1,1)) %>%
  
  
  #add indicator variable for unreliable estimates
  mutate(percent_relativese=percent_se/percent) %>%
  mutate(keep=if_else(percent_relativese>.3,0,1)) %>%
  
  #remove unneeded variables
  select(-c(count_sums,count_sums_se,
            denominator_overall,denominator_overall_se,percent_relativese)) %>%
  
  #add back in database variables
  mutate(week_start=all_of(week_minus3),
         week_end=all_of(week),
         timeframe='4-week rolling average',
         varname='freegroceriesorfreemeal',
         select_characteristics='Children in household')




####### -- 2C) Calculate Children Not Eating Enough Because Unaffordable -- #####

#start by importing the data needed from table 5
#sql code to import most recent 4 weeks of food 5 data
unaffordable_sql <- paste0("SELECT * FROM food5 WHERE state='",state_abbr,"' AND (week='",week,"' OR week='",week_minus1,"' OR week='",week_minus2,"' OR week='",week_minus3,"');")

unaffordable_4weeks_data <- dbGetQuery(con,unaffordable_sql) %>%
  
  #subset only for the category representing children not eating
  subset(categorynames=='childrennoteating') %>%
  #remove responses of Never and Did not report
  subset(select_characteristics != 'Never') %>%
  subset(select_characteristics !='Did not report') %>%
  
  #covert variables to numeric
  mutate(total=as.numeric(paste(total)),
         total_se=as.numeric(paste(total_se)),
         enough_not_types=as.numeric(paste(enough_not_types)),
         enough_not_types_se=as.numeric(paste(enough_not_types_se)),
         sometimes_not_enough=as.numeric(paste(sometimes_not_enough)),
         sometimes_not_enough_se=as.numeric(paste(sometimes_not_enough_se)),
         often_not_enough=as.numeric(paste(often_not_enough)),
         often_not_enough_se=as.numeric(paste(often_not_enough_se))) %>%
  
  #calculate the total minus did not report
  rowwise() %>%
  mutate(totalvalidresponses=sum(enough_not_types,sometimes_not_enough,often_not_enough,
                                 na.rm=TRUE),
         totalvalidresponses_se=sqrt((enough_not_types_se^2)+(sometimes_not_enough_se^2)+(often_not_enough_se^2))) %>%
  
  #combine total valid responses to get numerator across the four weeks
  group_by(categorynames,state) %>%
  summarise(children_not_eating=sum(totalvalidresponses,na.rm=TRUE),
            children_not_eating_se=sqrt(sum(totalvalidresponses_se^2)),.groups='keep')


#next, we need the total households with children from table food 3
#sql code to import most recent 4 weeks of food 3 data
totalhouseholds_sql <- paste0("SELECT * FROM food3b WHERE state='",state_abbr,"' AND (week='",week,"' OR week='",week_minus1,"' OR week='",week_minus2,"' OR week='",week_minus3,"');")

totalhouseholds_4weeks_data <- dbGetQuery(con,totalhouseholds_sql) %>%
  subset(select_characteristics=='Total') %>%
  
  #covert variables to numeric
  mutate(totalvalidresponses=as.numeric(paste(totalvalidresponses)),
         totalvalidresponses_se=as.numeric(paste(totalvalidresponses_se))) %>%
  
  #sum up the total valid responses
  group_by(state,categorynames) %>%
  summarise(totalhouseholdswithchildren=sum(totalvalidresponses,na.rm=TRUE),
            totalhouseholdswithchildren_se=sqrt(sum(totalvalidresponses_se^2)),.groups='keep') %>%
  ungroup %>%
  select(-c(categorynames))

#calculate the percentage
food_4weeks_data_unaffordable <- unaffordable_4weeks_data %>%
  left_join(totalhouseholds_4weeks_data,by=c('state'='state')) %>%
  
  #calculate percent
  mutate(percent=children_not_eating/totalhouseholdswithchildren,
         percent_se=(1/totalhouseholdswithchildren)*(sqrt((children_not_eating_se^2)-((percent^2)*(totalhouseholdswithchildren_se^2))))) %>%
  mutate(percent_se=if_else(is.na(percent_se),(1/totalhouseholdswithchildren)*(sqrt((children_not_eating_se^2)+((percent^2)*(totalhouseholdswithchildren_se^2)))),percent_se)) %>%
  mutate(percent_lci=percent-(percent_se*1.645)) %>%
  mutate(percent_uci=percent+(percent_se*1.645)) %>%
  
  #lower confidence interval shouldn't be less than zero, and upper more than 100
  mutate(percent_lci=replace(percent_lci,percent_lci<0,0)) %>%
  mutate(percent_uci=replace(percent_uci,percent_uci>1,1)) %>%
  
  #add indicator variable for unreliable estimates
  mutate(percent_relativese=percent_se/percent) %>%
  mutate(keep=if_else(percent_relativese>.3,0,1)) %>%
  
  #remove unneeded variables
  select(-c(children_not_eating,children_not_eating_se,
            totalhouseholdswithchildren,totalhouseholdswithchildren_se,percent_relativese)) %>%
  
  #add back in database variables
  mutate(week_start=all_of(week_minus3),
         week_end=all_of(week),
         timeframe='4-week rolling average',
         varname='childrennoteatingbecauseunaffordable',
         select_characteristics='Children in household')
  
  
  

  
####### -- 2D) COMBINE DATA TABLES -- #####
#combine three data tables
food_4weeks_combined <- food_4weeks_data %>%
  bind_rows(food_4weeks_data_overallfree2) %>%
  bind_rows(food_4weeks_data_childfree2) %>%
  bind_rows(food_4weeks_data_unaffordable)

#DATA QUALITY CHECKS
View(food_4weeks_combined)
```


## STEP 3: ADD ROLLING AVERAGE DATA TO SQL DATABASE
```{r}
dbWriteTable(con,'calculatedpercents',food_4weeks_combined,append=TRUE,row.names=FALSE)
```


## STEP 4: CALCULATE 12 "WEEK" AVERAGES FOR DEMOGRAPHICS
```{r}
week_minus1 <- week-1
week_minus2 <- week-2
week_minus3 <- week-3
week_minus4 <- week-4
week_minus5 <- week-5
week_minus6 <- week-6
week_minus7 <- week-7
week_minus8 <- week-8
week_minus9 <- week-9
week_minus10 <- week-10
week_minus11 <- week-11

####### -- 2A) Calculate the Sometimes/Often Not Enough Percentages -- #####

#sql code to import most recent 12 weeks of food 2b data
food_sql <- paste0("SELECT * FROM food2b WHERE state='",state_abbr,"' AND (week='",week,"' OR week='",week_minus1,"' OR week='",week_minus2,"' OR week='",week_minus3,"' OR week='",week_minus4,"' OR week='",week_minus5,"' OR week='",week_minus6,"' OR week='",week_minus7,"' OR week='",week_minus8,"' OR week='",week_minus9,"' OR week='",week_minus10,"' OR week='",week_minus11,"');")


food_12weeks_data <- dbGetQuery(con,food_sql) %>%
  
  subset(categorynames=='age' | categorynames=='race' | categorynames=='total' 
         | categorynames=='income') %>%
  
  #combine racial groups and income groups
  mutate(select_characteristics=replace(select_characteristics,
                                        select_characteristics=='Two or more races + Other races, not Hispanic' | 
                                          select_characteristics=='Black alone, not Hispanic' | 
                                          select_characteristics=='Asian alone, not Hispanic','Other/multiracial, not Latino')) %>%
  mutate(select_characteristics=replace(select_characteristics,
                                        select_characteristics=='Less than $25,000' | 
                                          select_characteristics=='$25,000 - $34,999' | 
                                          select_characteristics=='$35,000 - $49,999','Less than $50,000')) %>%
  mutate(select_characteristics=replace(select_characteristics,
                                        select_characteristics=='$50,000 - $74,999' | 
                                          select_characteristics=='$75,000 - $99,999' | 
                                          select_characteristics=='$100,000 - $149,999'
                                         | select_characteristics=='$150,000 - $199,999' | select_characteristics=='$200,000 and above','More than $50,000')) %>%
  subset(select_characteristics != 'Did not report') %>%
  
  #covert variables to numeric
  mutate(foodinsufficient_numerator=as.numeric(paste(foodinsufficient_numerator)),
         foodinsufficient_numerator_se=as.numeric(paste(foodinsufficient_numerator_se)),
         totalvalidresponses=as.numeric(paste(totalvalidresponses)),
         totalvalidresponses_se=as.numeric(paste(totalvalidresponses_se))) %>%
  
  #combine 12 weeks of data
  group_by(select_characteristics,categorynames,state) %>%
  summarise(foodinsufficient_numerator=sum(foodinsufficient_numerator,na.rm=TRUE),
            foodinsufficient_numerator_se=sqrt((sum(foodinsufficient_numerator_se^2))),
            totalvalidresponses=sum(totalvalidresponses,na.rm=TRUE),
            totalvalidresponses_se=sqrt((sum(totalvalidresponses_se^2))),
            .groups='keep') %>%
  ungroup %>%
  arrange(categorynames,select_characteristics) %>%
  
  #calculate percent food insufficient, se, and confidence interval
  mutate(percent=foodinsufficient_numerator/totalvalidresponses) %>%
  mutate(percent_se=(1/totalvalidresponses)*(sqrt((foodinsufficient_numerator_se^2)-((percent^2)*(totalvalidresponses_se^2))))) %>%
  mutate(percent_se=if_else(is.na(percent_se),(1/totalvalidresponses)*(sqrt((foodinsufficient_numerator_se^2)+((percent^2)*(totalvalidresponses_se^2)))),percent_se)) %>%
  mutate(percent_lci=percent-(percent_se*1.645)) %>%
  mutate(percent_uci=percent+(percent_se*1.645)) %>%
  
  #lower confidence interval shouldn't be less than zero, and upper more than 100
  mutate(percent_lci=replace(percent_lci,percent_lci<0,0)) %>%
  mutate(percent_uci=replace(percent_uci,percent_uci>1,1)) %>%
  
  
  #add indicator variable for unreliable estimates
  mutate(percent_relativese=percent_se/percent) %>%
  mutate(keep=if_else(percent_relativese>.3,0,1)) %>%
  
  #remove unneeded variables
  select(-c(foodinsufficient_numerator,foodinsufficient_numerator_se,
            totalvalidresponses,totalvalidresponses_se,percent_relativese)) %>%
  
  #rename some characteristics:
  mutate(select_characteristics=replace(select_characteristics,select_characteristics=='Total','All adults')) %>%
  mutate(select_characteristics=replace(select_characteristics,select_characteristics=='	
Hispanic or Latino (may be of any race)','Latino (any race)')) %>%
  mutate(select_characteristics=replace(select_characteristics,select_characteristics=='White alone, not Hispanic','White, not Latino')) %>%
  
  #add back in database variables
  mutate(week_start=all_of(week_minus11),
         week_end=all_of(week),
         timeframe='12-week rolling average',
         varname='sometimesoftennotenoughtoeat')




#DATA QUALITY CHECKS
View(food_12weeks_data)
```


## STEP 5: ADD ROLLING AVERAGE DATA TO SQL DATABASE
```{r}
dbWriteTable(con,'calculatedpercents',food_12weeks_data,append=TRUE,row.names=FALSE)
```



