---
title: "Household Pulse Health 2a"
output: html_document
---

#THIS CODE WORKS FOR WEEKS 28-PRESENT
#IMPORT HOUSEHOLD PULSE DATA FROM TABLE HEALTH 2a
#STORE DATA IN SQL DATABASE: health2a

## This code simply imports the data from Health Table2a.


```{r}
library(tidyverse)
library(readxl)
library(httr)
library(stringr)
```

## STEP 1: UPDATE THE MOST RECENT WEEK FOR DOWNLOADING DATA
```{r}
week <- 33
state_abbr <- 'SD'
```

## STEP 2: IMPORT MOST RECENT WEEK OF DATA
```{r}
#Parent directory by week: https://www2.census.gov/programs-surveys/demo/tables/hhp/2021/

#Health table 2a column names
health2a_colnames <-
  c("select_characteristics","nervous_not_at_all","nervous_several_days","nervous_half","nervous_everyday","nervous_did_not_report","worry_not_at_all","worry_several_days","worry_half","worry_everyday","worry_did_not_report")
categorynames <- c('total','age','age','age','age','age','age','age',
                          'sex','sex',
                          'race','race','race','race','race',
                          'education','education','education','education',
                          'maritalstatus','maritalstatus','maritalstatus',
                          'maritalstatus','maritalstatus',
                          'householdsize','householdsize','householdsize',
                          'householdsize','householdsize','householdsize',
                          'householdsize','children','children','lossincome',
                          'lossincome','lossincome',
                          'currentemploy','currentemploy',
                          'currentemploy',
                          'income','income','income','income','income','income',
                          'income','income','income',
                          'spending','spending','spending',
                          'spending','spending','spending',
                          'spending','spending','spending','spending',
                          'military','military','military','military',
                          'military','military',
                          'seeing','seeing','seeing','seeing','seeing',
                          'hearing','hearing','hearing','hearing','hearing',
                          'concentrating','concentrating','concentrating',
                          'concentrating','concentrating',
                          'walking','walking','walking','walking','walking')

#Import from Census parent directory
temp.file <- paste(tempfile(),".xlsx",sep = "")
download.file(paste0
                ("https://www2.census.gov/programs-surveys/demo/tables/hhp/2021/wk",week,"/health2a_week",week,".xlsx"), 
                temp.file, mode = "wb")

health2a_data <- read_excel(temp.file, sheet=paste0(state_abbr),skip = 6) %>%
    set_names(health2a_colnames) %>%
  
  #remove extra rows at beginning and end
    subset(!is.na(select_characteristics)) %>%
    filter(!str_detect(select_characteristics, "Totals may not sum")) %>%
    filter(!str_detect(select_characteristics, "The Census Bureau considers")) %>%
    filter(!str_detect(select_characteristics, "Select characteristics")) %>%
  
  #remove row headers, but then add back in category name
    subset(!is.na(nervous_not_at_all)) %>%
    add_column(categorynames) %>%
  
  #assign dashes to NA
    na_if("-") %>%
  
  #convert numeric variables to numeric
    mutate_at(vars(nervous_not_at_all,nervous_several_days,nervous_half,nervous_everyday,nervous_did_not_report,worry_not_at_all,worry_several_days,worry_half,worry_everyday,worry_did_not_report),
              list(~ as.numeric(.))) %>%
  
  #add in variable for week and state
  mutate(week=all_of(week)) %>%
  mutate(state=all_of(state_abbr)) %>%
  
  #calculate numerator, denominator %>%
  rowwise() %>%
  #calculate non-response rate for good measure
  mutate(nervous_nonresponse_percent=nervous_did_not_report/(nervous_not_at_all+nervous_several_days+nervous_half+nervous_everyday+nervous_did_not_report),
         worry_nonresponse_percent=worry_did_not_report/(worry_not_at_all+worry_several_days+worry_half+worry_everyday+worry_did_not_report))

##DATA QUALITY CHECKS
View(health2a_data)

```

## STEP 3: IMPORT AND ADD STANDARD ERROR
```{r}
#Parent directory by week: https://www2.census.gov/programs-surveys/demo/tables/hhp/2021/

#Import from Census parent directory
temp.file <- paste(tempfile(),".xlsx",sep = "")
download.file(paste0
                ("https://www2.census.gov/programs-surveys/demo/tables/hhp/2021/wk",week,"/health2a_se_week",week,".xlsx"), 
                temp.file, mode = "wb")

health2a_data_se <- read_excel(temp.file, sheet=paste0(state_abbr),skip = 6) %>%
    set_names(health2a_colnames) %>%
  
  #remove extra rows at beginning and end
    subset(!is.na(select_characteristics)) %>%
    filter(!str_detect(select_characteristics, "Totals may not sum")) %>%
    filter(!str_detect(select_characteristics, "The Census Bureau considers")) %>%
    filter(!str_detect(select_characteristics, "Select characteristics")) %>%
  
  #remove row headers, but then add back in category name
    subset(!is.na(nervous_not_at_all)) %>%
    add_column(categorynames) %>%
  
  #assign dashes to na
    na_if("-") %>%
  
  #convert numeric variables to numeric
    mutate_at(vars(nervous_not_at_all,nervous_several_days,nervous_half,nervous_everyday,nervous_did_not_report,worry_not_at_all,worry_several_days,worry_half,worry_everyday,worry_did_not_report),
              list(~ as.numeric(.))) %>%
  
  #convert NA's to zero
  replace(is.na(.),0) %>%
  
  #add in variable for week and state
  mutate(week=all_of(week)) %>%
  mutate(state=all_of(state_abbr)) %>%
  
  #change variable names to _se
  rename(nervous_not_at_all_se=nervous_not_at_all,
         nervous_several_days_se=nervous_several_days,
         nervous_half_se=nervous_half,
         nervous_everyday_se=nervous_everyday,
         nervous_did_not_report_se=nervous_did_not_report,
         worry_not_at_all_se=worry_not_at_all,
         worry_several_days_se=worry_several_days,
         worry_half_se=worry_half,
         worry_everyday_se=worry_everyday,
         worry_did_not_report_se=worry_did_not_report) 


##DATA QUALITY CHECKS
View(health2a_data_se)

```

## STEP 4: COMBINE DATA AND STANDARD ERROR
```{r}
health2a_datacombined <- health2a_data %>%
  left_join(health2a_data_se,by=c('select_characteristics'='select_characteristics',
                             'week'='week','categorynames'='categorynames',
                             'state'='state'))

##DATA QUALITY CHECKS
View(health2a_datacombined)
```

## STEP 5: ADD MOST RECENT WEEK DATA TO SQL DATABASE
```{r}
dbWriteTable(con,'health2a',health2a_datacombined,append=TRUE,row.names=FALSE)
```

