---
title: "Household Pulse Health 2b"
output: html_document
---

#THIS CODE WORKS FOR WEEKS 22-PRESENT
#IMPORT HOUSEHOLD PULSE DATA FROM TABLE HEALTH 2b
#STORE DATA IN SQL DATABASE: health2b

## This code simply imports the data from Health Table2b.


```{r}
library(tidyverse)
library(readxl)
library(httr)
library(stringr)
```

## STEP 1: UPDATE THE MOST RECENT WEEK FOR DOWNLOADING DATA
```{r}
week <- 27
state_abbr <- 'SD'
```

## STEP 2: IMPORT MOST RECENT WEEK OF DATA
```{r}
#Parent directory by week: https://www2.census.gov/programs-surveys/demo/tables/hhp/2021/

#Health table 2b column names
health2b_colnames <-
  c("select_characteristics","nopleasure_not_at_all","nopleasure_several_days","nopleasure_half","nopleasure_everyday","nopleasure_did_not_report","depressed_not_at_all","depressed_several_days","depressed_half","depressed_everyday","depressed_did_not_report")
categorynames <- c('total','age','age','age','age','age','age','age',
                          'sex','sex',
                          'race','race','race','race','race',
                          'education','education','education','education',
                          'maritalstatus','maritalstatus','maritalstatus',
                          'maritalstatus','maritalstatus',
                          'householdsize','householdsize','householdsize',
                          'householdsize','householdsize','householdsize',
                          'householdsize','children','children','lossincome',
                          'lossincome','lossincome',
                          'currentemploy','currentemploy',
                          'currentemploy',
                          'income','income','income','income','income','income',
                          'income','income','income',
                          'spending','spending','spending',
                          'spending','spending','spending',
                          'spending','spending','spending')

#Import from Census parent directory
temp.file <- paste(tempfile(),".xlsx",sep = "")
download.file(paste0
                ("https://www2.census.gov/programs-surveys/demo/tables/hhp/2021/wk",week,"/health2b_week",week,".xlsx"), 
                temp.file, mode = "wb")

health2b_data <- read_excel(temp.file, sheet=paste0(state_abbr),skip = 6) %>%
    set_names(health2b_colnames) %>%
  
  #remove extra rows at beginning and end
    subset(!is.na(select_characteristics)) %>%
    filter(!str_detect(select_characteristics, "Totals may not sum")) %>%
    filter(!str_detect(select_characteristics, "The Census Bureau considers")) %>%
    filter(!str_detect(select_characteristics, "Select characteristics")) %>%
  
  #remove row headers, but then add back in category name
    subset(!is.na(nopleasure_not_at_all)) %>%
    add_column(categorynames) %>%
  
  #assign dashes to NA
    na_if("-") %>%
  
  #convert numeric variables to numeric
    mutate_at(vars(nopleasure_not_at_all,nopleasure_several_days,nopleasure_half,nopleasure_everyday,nopleasure_did_not_report,depressed_not_at_all,depressed_several_days,depressed_half,depressed_everyday,depressed_did_not_report),
              list(~ as.numeric(.))) %>%
  
  #add in variable for week and state
  mutate(week=all_of(week)) %>%
  mutate(state=all_of(state_abbr)) %>%
  
  #calculate numerator, denominator %>%
  rowwise() %>%
  #calculate non-response rate for good measure
  mutate(nopleasure_nonresponse_percent=nopleasure_did_not_report/(nopleasure_not_at_all+nopleasure_several_days+nopleasure_half+nopleasure_everyday+nopleasure_did_not_report),
         depressed_nonresponse_percent=depressed_did_not_report/(depressed_not_at_all+depressed_several_days+depressed_half+depressed_everyday+depressed_did_not_report))

##DATA QUALITY CHECKS
View(health2b_data)

```

## STEP 3: IMPORT AND ADD STANDARD ERROR
```{r}
#Parent directory by week: https://www2.census.gov/programs-surveys/demo/tables/hhp/2021/

#Health table 2b column names
health2b_colnames <-
  c("select_characteristics","nopleasure_not_at_all","nopleasure_several_days","nopleasure_half","nopleasure_everyday","nopleasure_did_not_report","depressed_not_at_all","depressed_several_days","depressed_half","depressed_everyday","depressed_did_not_report")
categorynames <- c('total','age','age','age','age','age','age','age',
                          'sex','sex',
                          'race','race','race','race','race',
                          'education','education','education','education',
                          'maritalstatus','maritalstatus','maritalstatus',
                          'maritalstatus','maritalstatus',
                          'householdsize','householdsize','householdsize',
                          'householdsize','householdsize','householdsize',
                          'householdsize','children','children','lossincome',
                          'lossincome','lossincome',
                          'currentemploy','currentemploy',
                          'currentemploy',
                          'income','income','income','income','income','income',
                          'income','income','income',
                          'spending','spending','spending',
                          'spending','spending','spending',
                          'spending','spending','spending')

#Import from Census parent directory
temp.file <- paste(tempfile(),".xlsx",sep = "")
download.file(paste0
                ("https://www2.census.gov/programs-surveys/demo/tables/hhp/2021/wk",week,"/health2b_se_week",week,".xlsx"), 
                temp.file, mode = "wb")

health2b_data_se <- read_excel(temp.file, sheet=paste0(state_abbr),skip = 6) %>%
    set_names(health2b_colnames) %>%
  
  #remove extra rows at beginning and end
    subset(!is.na(select_characteristics)) %>%
    filter(!str_detect(select_characteristics, "Totals may not sum")) %>%
    filter(!str_detect(select_characteristics, "The Census Bureau considers")) %>%
    filter(!str_detect(select_characteristics, "Select characteristics")) %>%
  
  #remove row headers, but then add back in category name
    subset(!is.na(nopleasure_not_at_all)) %>%
    add_column(categorynames) %>%
  
  #assign dashes to na
    na_if("-") %>%
  
  #convert numeric variables to numeric
    mutate_at(vars(nopleasure_not_at_all,nopleasure_several_days,nopleasure_half,nopleasure_everyday,nopleasure_did_not_report,depressed_not_at_all,depressed_several_days,depressed_half,depressed_everyday,depressed_did_not_report),
              list(~ as.numeric(.))) %>%
  
  #convert NA's to zero
  replace(is.na(.),0) %>%
  
  #add in variable for week and state
  mutate(week=all_of(week)) %>%
  mutate(state=all_of(state_abbr)) %>%
  
  #change variable names to _se
  rename(nopleasure_not_at_all_se=nopleasure_not_at_all,
         nopleasure_several_days_se=nopleasure_several_days,
         nopleasure_half_se=nopleasure_half,
         nopleasure_everyday_se=nopleasure_everyday,
         nopleasure_did_not_report_se=nopleasure_did_not_report,
         depressed_not_at_all_se=depressed_not_at_all,
         depressed_several_days_se=depressed_several_days,
         depressed_half_se=depressed_half,
         depressed_everyday_se=depressed_everyday,
         depressed_did_not_report_se=depressed_did_not_report) 


##DATA QUALITY CHECKS
View(health2b_data_se)

```

## STEP 4: COMBINE DATA AND STANDARD ERROR
```{r}
health2b_datacombined <- health2b_data %>%
  left_join(health2b_data_se,by=c('select_characteristics'='select_characteristics',
                             'week'='week','categorynames'='categorynames',
                             'state'='state'))

##DATA QUALITY CHECKS
View(health2b_datacombined)
```

## STEP 5: ADD MOST RECENT WEEK DATA TO SQL DATABASE
```{r}
dbWriteTable(con,'health2b',health2b_datacombined,append=TRUE,row.names=FALSE)
```

