---
title: "Household Pulse Health 1"
output: html_document
---

#THIS CODE WORKS FOR WEEKS 13-21
#IMPORT HOUSEHOLD PULSE DATA FROM TABLE HEALTH 1
#STORE DATA IN SQL DATABASE: health1

## This code simply imports the data from Health Table1, and calculates the yes numerator, denominator, and a nonresponse percent.


```{r}
library(tidyverse)
library(readxl)
library(httr)
library(stringr)
```

## STEP 1: UPDATE THE MOST RECENT WEEK FOR DOWNLOADING DATA
```{r}
week <- 21
state_abbr <- 'SD'
```

## STEP 2: IMPORT MOST RECENT WEEK OF DATA
```{r}
#Parent directory by week: https://www2.census.gov/programs-surveys/demo/tables/hhp/2020/

#Health table 1 column names
health1_colnames <-
  c("select_characteristics","delaycare_yes", "delaycare_no", "delaycare_did_not_report", "unrelated_delaycare_yes", "unrelated_delaycare_no", "unrelated_delaycare_did_not_report")
categorynames <- c('total','age','age','age','age','age','age','age',
                          'sex','sex',
                          'race','race','race','race','race',
                          'education','education','education','education',
                          'maritalstatus','maritalstatus','maritalstatus',
                          'maritalstatus','maritalstatus',
                          'householdsize','householdsize','householdsize',
                          'householdsize','householdsize','householdsize',
                          'householdsize','children','children','lossincome',
                          'lossincome','lossincome',
                          'currentemploy','currentemploy',
                          'currentemploy',
                          'income','income','income','income','income','income',
                          'income','income','income',
                          'spending','spending','spending',
                          'spending','spending','spending',
                          'spending','spending','spending')

#Import from Census parent directory
temp.file <- paste(tempfile(),".xlsx",sep = "")
download.file(paste0
                ("https://www2.census.gov/programs-surveys/demo/tables/hhp/2020/wk",week,"/health1_week",week,".xlsx"), 
                temp.file, mode = "wb")

health1_data <- read_excel(temp.file, sheet=paste0(state_abbr),skip = 6) %>%
    set_names(health1_colnames) %>%
  
  #remove extra rows at beginning and end
    subset(!is.na(select_characteristics)) %>%
    filter(!str_detect(select_characteristics, "Totals may not sum")) %>%
    filter(!str_detect(select_characteristics, "The Census Bureau considers")) %>%
    filter(!str_detect(select_characteristics, "Select characteristics")) %>%
  
  #remove row headers, but then add back in category name
    subset(!is.na(delaycare_yes)) %>%
    add_column(categorynames) %>%
  
  #assign dashes to NA
    na_if("-") %>%
  
  #convert numeric variables to numeric
    mutate_at(vars(delaycare_yes,delaycare_no,delaycare_did_not_report,unrelated_delaycare_yes,unrelated_delaycare_no,unrelated_delaycare_did_not_report),
              list(~ as.numeric(.))) %>%
  
  #add in variable for week and state
  mutate(week=all_of(week)) %>%
  mutate(state=all_of(state_abbr)) %>%
  
  #calculate numerator, denominator %>%
  rowwise() %>%
  mutate(unrelated_delaycare_numerator=sum(unrelated_delaycare_yes,na.rm=TRUE),
         unrelated_delaycare_denominator=sum(unrelated_delaycare_yes,unrelated_delaycare_no,na.rm=TRUE)) %>%
  #calculate non-response rate for good measure
  mutate(unrelated_delaycare_nonresponse_percent=unrelated_delaycare_did_not_report/(unrelated_delaycare_denominator+unrelated_delaycare_did_not_report))

##DATA QUALITY CHECKS
View(health1_data)

```

## STEP 3: IMPORT AND ADD STANDARD ERROR
```{r}
#Parent directory by week: https://www2.census.gov/programs-surveys/demo/tables/hhp/2020/

#Health table 1 column names
health1_colnames <-
  c("select_characteristics","delaycare_yes", "delaycare_no", "delaycare_did_not_report", "unrelated_delaycare_yes", "unrelated_delaycare_no", "unrelated_delaycare_did_not_report")
categorynames <- c('total','age','age','age','age','age','age','age',
                          'sex','sex',
                          'race','race','race','race','race',
                          'education','education','education','education',
                          'maritalstatus','maritalstatus','maritalstatus',
                          'maritalstatus','maritalstatus',
                          'householdsize','householdsize','householdsize',
                          'householdsize','householdsize','householdsize',
                          'householdsize','children','children','lossincome',
                          'lossincome','lossincome',
                          'currentemploy','currentemploy',
                          'currentemploy',
                          'income','income','income','income','income','income',
                          'income','income','income',
                          'spending','spending','spending',
                          'spending','spending','spending',
                          'spending','spending','spending')

#Import from Census parent directory
temp.file <- paste(tempfile(),".xlsx",sep = "")
download.file(paste0
                ("https://www2.census.gov/programs-surveys/demo/tables/hhp/2020/wk",week,"/health1_se_week",week,".xlsx"), 
                temp.file, mode = "wb")

health1_data_se <- read_excel(temp.file, sheet=paste0(state_abbr),skip = 6) %>%
    set_names(health1_colnames) %>%
  
  #remove extra rows at beginning and end
    subset(!is.na(select_characteristics)) %>%
    filter(!str_detect(select_characteristics, "Totals may not sum")) %>%
    filter(!str_detect(select_characteristics, "The Census Bureau considers")) %>%
    filter(!str_detect(select_characteristics, "Select characteristics")) %>%
  
  #remove row headers, but then add back in category name
    subset(!is.na(delaycare_yes)) %>%
    add_column(categorynames) %>%
  
  #assign dashes to na
    na_if("-") %>%
  
  #convert numeric variables to numeric
    mutate_at(vars(delaycare_yes,delaycare_no,delaycare_did_not_report,unrelated_delaycare_yes,unrelated_delaycare_no,unrelated_delaycare_did_not_report),
              list(~ as.numeric(.))) %>%
  
  #convert NA's to zero
  replace(is.na(.),0) %>%
  
  #add in variable for week and state
  mutate(week=all_of(week)) %>%
  mutate(state=all_of(state_abbr)) %>%
  
  #change variable names to _se
  rename(delaycare_yes_se=delaycare_yes,
         delaycare_no_se=delaycare_no,
         delaycare_did_not_report_se=delaycare_did_not_report,
         unrelated_delaycare_yes_se=unrelated_delaycare_yes,
         unrelated_delaycare_no_se=unrelated_delaycare_no,
         unrelated_delaycare_did_not_report_se=unrelated_delaycare_did_not_report) %>%
  
  #calculate numerator, denominator se's
  mutate(unrelated_delaycare_numerator_se=unrelated_delaycare_yes_se,
         unrelated_delaycare_denominator_se=sqrt((unrelated_delaycare_yes_se^2)+(unrelated_delaycare_no_se^2))) 

##DATA QUALITY CHECKS
View(health1_data_se)

```

## STEP 4: COMBINE DATA AND STANDARD ERROR
```{r}
health1_datacombined <- health1_data %>%
  left_join(health1_data_se,by=c('select_characteristics'='select_characteristics',
                             'week'='week','categorynames'='categorynames',
                             'state'='state'))

##DATA QUALITY CHECKS
View(health1_datacombined)
```

## STEP 5: ADD MOST RECENT WEEK DATA TO SQL DATABASE
```{r}
dbWriteTable(con,'health1',health1_datacombined,append=TRUE,row.names=FALSE)
```

