---
title: "Household Pulse Housing 1B"
output: html_document
---

#IMPORT HOUSEHOLD PULSE DATA FROM TABLE HOUSING 1B
#STORE DATA IN SQL DATABASE: housing1b

## This code simply imports the data from Housing Table 1b, and calculates the not caught up numerator, denominator, and a nonresponse percent.


```{r}
library(tidyverse)
library(readxl)
library(httr)
library(stringr)
```

## STEP 1: UPDATE THE MOST RECENT WEEK FOR DOWNLOADING DATA
```{r}
week <- 29
state_abbr <- 'SD'
```

## STEP 2: IMPORT MOST RECENT WEEK OF DATA
```{r}
#Parent directory by week: https://www2.census.gov/programs-surveys/demo/tables/hhp/2021/

#Housing table 1b column names
housing1b_colnames <-
  c("select_characteristics", "total", "occupied_without_rent", "yes_caughtup", "no_notcaughtup", "did_not_report", "did_not_report_tenure")
categorynames <- c('total','age','age','age','age','age',
                          'sex','sex',
                          'race','race','race','race','race',
                          'education','education','education','education',
                          'maritalstatus','maritalstatus','maritalstatus',
                          'maritalstatus','maritalstatus',
                          'householdsize','householdsize','householdsize',
                          'householdsize','householdsize','householdsize',
                          'householdsize','children','children','lossincome',
                          'lossincome','lossincome',
                          'currentemploy','currentemploy',
                          'currentemploy',
                          'income','income','income','income','income','income',
                          'income','income','income',
                          'spending','spending','spending',
                          'spending','spending','spending',
                          'spending','spending','spending','spending',
                          'military','military','military','military',
                          'military','military',
                          'seeing','seeing','seeing','seeing','seeing',
                          'hearing','hearing','hearing','hearing','hearing',
                          'concentrating','concentrating','concentrating',
                          'concentrating','concentrating',
                          'walking','walking','walking','walking','walking')

#Import from Census parent directory
temp.file <- paste(tempfile(),".xlsx",sep = "")
download.file(paste0
                ("https://www2.census.gov/programs-surveys/demo/tables/hhp/2021/wk",week,"/housing1b_week",week,".xlsx"), 
                temp.file, mode = "wb")

housing1b_data <- read_excel(temp.file, sheet=paste0(state_abbr),skip = 5) %>%
    set_names(housing1b_colnames) %>%
  
  #remove extra rows at beginning and end
    subset(!is.na(select_characteristics)) %>%
    filter(!str_detect(select_characteristics, "Totals may not sum")) %>%
    filter(!str_detect(select_characteristics, "The Census Bureau considers")) %>%
    filter(!str_detect(select_characteristics, "Select characteristics")) %>%
  
  #remove row headers, but then add back in category name
    subset(!is.na(total)) %>%
    add_column(categorynames) %>%
  
  #assign dashes to NA
    na_if("-") %>%
  
  #convert numeric variables to numeric
    mutate_at(vars(total,occupied_without_rent,yes_caughtup,
                   no_notcaughtup,did_not_report,did_not_report_tenure),
              list(~ as.numeric(.))) %>%
  
  #add in variable for week and state
  mutate(week=all_of(week)) %>%
  mutate(state=all_of(state_abbr)) %>%
  
  #calculate numerator, denominator %>%
  rowwise() %>%
  mutate(notcaughtup_numerator=sum(no_notcaughtup,na.rm=TRUE),
         notcaughtup_denominator=sum(yes_caughtup,no_notcaughtup,na.rm=TRUE)) %>%
  #calculate non-response rate for good measure
  mutate(nonresponse_percent=did_not_report/total)

##DATA QUALITY CHECKS
View(housing1b_data)

```

## STEP 3: IMPORT AND ADD STANDARD ERROR
```{r}
#Parent directory by week: https://www2.census.gov/programs-surveys/demo/tables/hhp/2021/

#Import from Census parent directory
temp.file <- paste(tempfile(),".xlsx",sep = "")
download.file(paste0
                ("https://www2.census.gov/programs-surveys/demo/tables/hhp/2021/wk",week,"/housing1b_se_week",week,".xlsx"), 
                temp.file, mode = "wb")

housing1b_data_se <- read_excel(temp.file, sheet=paste0(state_abbr),skip = 5) %>%
    set_names(housing1b_colnames) %>%
  
  #remove extra rows at beginning and end
    subset(!is.na(select_characteristics)) %>%
    filter(!str_detect(select_characteristics, "Totals may not sum")) %>%
    filter(!str_detect(select_characteristics, "The Census Bureau considers")) %>%
    filter(!str_detect(select_characteristics, "Select characteristics")) %>%
  
  #remove row headers, but then add back in category name
    subset(!is.na(total)) %>%
    add_column(categorynames) %>%
  
  #assign dashes to na
    na_if("-") %>%
  
  #convert numeric variables to numeric
    mutate_at(vars(total,occupied_without_rent,yes_caughtup,
                   no_notcaughtup,did_not_report,did_not_report_tenure),
              list(~ as.numeric(.))) %>%
  
  #convert NA's to zero
  replace(is.na(.),0) %>%
  
  #add in variable for week and state
  mutate(week=all_of(week)) %>%
  mutate(state=all_of(state_abbr)) %>%
  
  #change variable names to _se
  rename(total_se=total,
         occupied_without_rent_se=occupied_without_rent,
         yes_caughtup_se=yes_caughtup,
         no_notcaughtup_se=no_notcaughtup,
         did_not_report_se=did_not_report,
         did_not_report_tenure_se=did_not_report_tenure) %>%
  
  #calculate numerator, denominator se's
  mutate(notcaughtup_numerator_se=no_notcaughtup_se,
         notcaughtup_denominator_se=sqrt((yes_caughtup_se^2)+(no_notcaughtup_se^2))) 

##DATA QUALITY CHECKS
View(housing1b_data_se)

```

## STEP 4: COMBINE DATA AND STANDARD ERROR
```{r}
housing1b_datacombined <- housing1b_data %>%
  left_join(housing1b_data_se,by=c('select_characteristics'='select_characteristics',
                             'week'='week','categorynames'='categorynames',
                             'state'='state'))

##DATA QUALITY CHECKS
View(housing1b_datacombined)
```

## STEP 5: ADD MOST RECENT WEEK DATA TO SQL DATABASE
```{r}
dbWriteTable(con,'housing1b',housing1b_datacombined,append=TRUE,row.names=FALSE)
```

