---
title: "Household Pulse Food Calculations"
output: html_document
---

## This code uses data imported from Housing Table 1b in order to calculate percentages used in the data viz.



```{r}
library(tidyverse)
library(readxl)
library(httr)
library(stringr)
```

## STEP 1: UPDATE THE MOST RECENT WEEK FOR DOWNLOADING DATA
```{r}
week <- 25
state_abbr <- 'SD'
```


## STEP 2: CALCULATE ROLLING 4 "WEEK" AVERAGES
# 2a) For Percent Not Caught Up on Rent
```{r}
week_minus1 <- week-1
week_minus2 <- week-2
week_minus3 <- week-3

####### -- 2A) Calculate the Percent Not Caught up on Rent -- #####

#sql code to import most recent 4 weeks of food 2b data
housing_sql <- paste0("SELECT * FROM housing1b WHERE state='",state_abbr,"' AND (week='",week,"' OR week='",week_minus1,"' OR week='",week_minus2,"' OR week='",week_minus3,"');")

housing_4weeks_data <- dbGetQuery(con,housing_sql) %>%
  
  subset(categorynames=='children' | categorynames=='total') %>%
  
  #covert variables to numeric
  mutate(notcaughtup_numerator=as.numeric(paste(notcaughtup_numerator)),
         notcaughtup_numerator_se=as.numeric(paste(notcaughtup_numerator_se)),
         notcaughtup_denominator=as.numeric(paste(notcaughtup_denominator)),
         notcaughtup_denominator_se=as.numeric(paste(notcaughtup_denominator_se))) %>%
  
  #combine four weeks of data
  group_by(select_characteristics,categorynames,state) %>%
  summarise(notcaughtup_numerator=sum(notcaughtup_numerator,na.rm=TRUE),
            notcaughtup_numerator_se=sqrt((sum(notcaughtup_numerator_se^2))),
            notcaughtup_denominator=sum(notcaughtup_denominator,na.rm=TRUE),
            notcaughtup_denominator_se=sqrt((sum(notcaughtup_denominator_se^2))),
            .groups='keep') %>%
  ungroup %>%
  arrange(categorynames,select_characteristics) %>%
  
  #calculate percent food insufficient, se, and confidence interval
  mutate(percent=notcaughtup_numerator/notcaughtup_denominator) %>%
  mutate(percent_se=(1/notcaughtup_denominator)*(sqrt((notcaughtup_numerator_se^2)-((percent^2)*(notcaughtup_denominator_se^2))))) %>%
  mutate(percent_se=if_else(is.na(percent_se),(1/notcaughtup_denominator)*(sqrt((notcaughtup_numerator_se^2)+((percent^2)*(notcaughtup_denominator_se^2)))),percent_se)) %>%
  mutate(percent_lci=percent-(percent_se*1.645)) %>%
  mutate(percent_uci=percent+(percent_se*1.645)) %>%
  
  #lower confidence interval shouldn't be less than zero, and upper more than 100
  mutate(percent_lci=replace(percent_lci,percent_lci<0,0)) %>%
  mutate(percent_uci=replace(percent_uci,percent_uci>1,1)) %>%
  
  
  #add indicator variable for unreliable estimates
  mutate(percent_relativese=percent_se/percent) %>%
  mutate(keep=if_else(percent_relativese>.3,0,1)) %>%
  
  #remove unneeded variables
  select(-c(notcaughtup_numerator,notcaughtup_numerator_se,
            notcaughtup_denominator,notcaughtup_denominator_se,percent_relativese)) %>%
  
  #add back in database variables
  mutate(week_start=all_of(week_minus3),
         week_end=all_of(week),
         timeframe='4-week rolling average',
         varname='notcaughtuponrent')


#DATA QUALITY CHECKS
View(housing_4weeks_data)
```


## STEP 3: ADD ROLLING AVERAGE DATA TO SQL DATABASE
```{r}
dbWriteTable(con,'calculatedpercents',housing_4weeks_data,append=TRUE,row.names=FALSE)
```


## STEP 4: CALCULATE 12 "WEEK" AVERAGES FOR DEMOGRAPHICS
```{r}
week_minus1 <- week-1
week_minus2 <- week-2
week_minus3 <- week-3
week_minus4 <- week-4
week_minus5 <- week-5
week_minus6 <- week-6
week_minus7 <- week-7
week_minus8 <- week-8
week_minus9 <- week-9
week_minus10 <- week-10
week_minus11 <- week-11

####### -- 4A) Calculate the Percent Not Caught up on Rent -- #####

#sql code to import most recent 12 weeks of housing1b data
housing_sql <- paste0("SELECT * FROM housing1b WHERE state='",state_abbr,"' AND (week='",week,"' OR week='",week_minus1,"' OR week='",week_minus2,"' OR week='",week_minus3,"' OR week='",week_minus4,"' OR week='",week_minus5,"' OR week='",week_minus6,"' OR week='",week_minus7,"' OR week='",week_minus8,"' OR week='",week_minus9,"' OR week='",week_minus10,"' OR week='",week_minus11,"');")


housing_12weeks_data <- dbGetQuery(con,housing_sql) %>%
  
  subset(categorynames=='age' | categorynames=='race' | categorynames=='total' 
         | categorynames=='income') %>%
  
  #combine racial groups and income groups
  mutate(select_characteristics=replace(select_characteristics,
                                        select_characteristics=='Two or more races + Other races, not Hispanic' | 
                                          select_characteristics=='Black alone, not Hispanic' | 
                                          select_characteristics=='Asian alone, not Hispanic','Other/multiracial, not Latino')) %>%
  mutate(select_characteristics=replace(select_characteristics,
                                        select_characteristics=='Less than $25,000' | 
                                          select_characteristics=='$25,000 - $34,999' | 
                                          select_characteristics=='$35,000 - $49,999','Less than $50,000')) %>%
  mutate(select_characteristics=replace(select_characteristics,
                                        select_characteristics=='$50,000 - $74,999' | 
                                          select_characteristics=='$75,000 - $99,999' | 
                                          select_characteristics=='$100,000 - $149,999'
                                         | select_characteristics=='$150,000 - $199,999' | select_characteristics=='$200,000 and above','More than $50,000')) %>%
  subset(select_characteristics != 'Did not report') %>%
  
  #covert variables to numeric
  mutate(notcaughtup_numerator=as.numeric(paste(notcaughtup_numerator)),
         notcaughtup_numerator_se=as.numeric(paste(notcaughtup_numerator_se)),
         notcaughtup_denominator=as.numeric(paste(notcaughtup_denominator)),
         notcaughtup_denominator_se=as.numeric(paste(notcaughtup_denominator_se))) %>%
  
  #combine 12 weeks of data
  group_by(select_characteristics,categorynames,state) %>%
  summarise(notcaughtup_numerator=sum(notcaughtup_numerator,na.rm=TRUE),
            notcaughtup_numerator_se=sqrt((sum(notcaughtup_numerator_se^2))),
            notcaughtup_denominator=sum(notcaughtup_denominator,na.rm=TRUE),
            notcaughtup_denominator_se=sqrt((sum(notcaughtup_denominator_se^2))),
            .groups='keep') %>%
  ungroup %>%
  arrange(categorynames,select_characteristics) %>%
  
  #calculate percent food insufficient, se, and confidence interval
  mutate(percent=notcaughtup_numerator/notcaughtup_denominator) %>%
  mutate(percent_se=(1/notcaughtup_denominator)*(sqrt((notcaughtup_numerator_se^2)-((percent^2)*(notcaughtup_denominator_se^2))))) %>%
  mutate(percent_se=if_else(is.na(percent_se),(1/notcaughtup_denominator)*(sqrt((notcaughtup_numerator_se^2)+((percent^2)*(notcaughtup_denominator_se^2)))),percent_se)) %>%
  mutate(percent_lci=percent-(percent_se*1.645)) %>%
  mutate(percent_uci=percent+(percent_se*1.645)) %>%
  
  #lower confidence interval shouldn't be less than zero, and upper more than 100
  mutate(percent_lci=replace(percent_lci,percent_lci<0,0)) %>%
  mutate(percent_uci=replace(percent_uci,percent_uci>1,1)) %>%
  
  
  #add indicator variable for unreliable estimates
  mutate(percent_relativese=percent_se/percent) %>%
  mutate(keep=if_else(percent_relativese>.3,0,1)) %>%
  
  #remove unneeded variables
  select(-c(notcaughtup_numerator,notcaughtup_numerator_se,
            notcaughtup_denominator,notcaughtup_denominator_se,percent_relativese)) %>%
  
  #rename some characteristics:
  mutate(select_characteristics=replace(select_characteristics,select_characteristics=='Total','All adults')) %>%
  mutate(select_characteristics=replace(select_characteristics,select_characteristics=='	
Hispanic or Latino (may be of any race)','Latino (any race)')) %>%
  mutate(select_characteristics=replace(select_characteristics,select_characteristics=='White alone, not Hispanic','White, not Latino')) %>%
  
  #add back in database variables
  mutate(week_start=all_of(week_minus11),
         week_end=all_of(week),
         timeframe='12-week rolling average',
         varname='notcaughtuponrent')




#DATA QUALITY CHECKS
View(housing_12weeks_data)
```


## STEP 5: ADD ROLLING AVERAGE DATA TO SQL DATABASE
```{r}
dbWriteTable(con,'calculatedpercents',housing_12weeks_data,append=TRUE,row.names=FALSE)
```



